---
title: "RNAseAlert assay 07: substrate"
author: "Jan Sklenar"
date: "04/05/2020"
output: html_document
editor_options: 
chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Efect of substrate batch on RNAseAlert assay 
- We are testing two concentrations 1ul and 10ul. (The former is the recommended level)
- Joe's substrate (brown tube) - very old
- New from the kit (50pmol per tube dry - dissolved in 60 ul of kit  buffer)
- New in Apr 2020 (brown tube) in water (2uM in 1ml)

## Protocol:
![Copy of the original Excel table.](data/protocol7.svg)

#### Original RNAaseAlert [kit manual](https://www.thermofisher.com/document-connect/document-connect.html?url=https%3A%2F%2Fassets.thermofisher.com%2FTFS-Assets%2FLSG%2Fmanuals%2Ffm_1964.pdf&title=Uk5hc2VBbGVydCZ0cmFkZTsgTGFiIFRlc3QgS2l0IEluc3RydWN0aW9uIE1hbnVhbA==)


```{r libraries, echo=FALSE, include=FALSE}
#library("plyr")
library("dplyr")
#library("tidyr")
library("ggplot2")
```

## Data preparation
### Read in the data and parse some sample names 
```{r data, echo=TRUE, include=TRUE}
## read in raw data
raw <- read.delim("data/200504_RNAaseAlert_substrate2020-05-04T145821.txt",stringsAsFactors = FALSE)
colnames(raw)

#' remove column
raw <- raw[,1:10]
#' change names
colnames(raw)[7]
colnames(raw)[7] <- "MeasTimeSec"

#' parse 'Sample' to obtain sample number 'SampleNo'
raw$Sample[1]
#substr(raw$Sample,1,regexpr(' ', raw$Sample)-1) 
raw$SampleNo <- substr(raw$Sample,1,regexpr(' ', raw$Sample)-1)

#' in the same way extract replicate number - 'RepNo'
#substr(raw$Sample,regexpr(' ', raw$Sample)+1, regexpr('/', raw$Sample)-1)
raw$RepNo <- substr(raw$Sample,regexpr(' ', raw$Sample)+1, regexpr('/', raw$Sample)-1)

#' remove white space from 'Well' sting
#substr(raw$Well, 2, regexpr('$', raw$Well))[1:5]
raw$Well<- substr(raw$Well, 2, regexpr('$', raw$Well))
#' View(raw) 
str(raw)
```

### Corrections
Not this time.

```{r corrections, include=FALSE, echo=FALSE, eval=FALSE}
## ## Condition 1 and 2 are the same thing. Create new SampleComb where 2 is replace with 1

raw$SampleComb <- raw$SampleNo 
raw[raw$SampleComb==2,"SampleComb"] <- 1

unique(raw$SampleNo)
unique(raw$SampleComb)
#' Which means the control with +RNAse has 8 replicates, unlike all the other conditions that has only 4. 
```



### Subset of data for plotting
```{r cleaned data}
## Selected (cleaned) data, only columns needed for plotting
#' Well, Sample, SampleNo, RepNo, MeasTimeSec, RFU
colnames(raw)
df <- raw[,c(3,6,7,8,11,12)] 
head(df) 
str(df)
```
## Plots
### All data
```{r plots}
#' all the data
#plot(x=raw$MeasTimeSec,y=raw$RFU)

#' colour labeled samples
ggplot(data = df) +
  aes(x = MeasTimeSec, y=RFU) +
  geom_point(aes(color=SampleNo))
```

### Adjustments (if needed)
Not this time. A limit on axis x is need. It will be added to plots later.

```{r final result, include=FALSE, echo=FALSE}
# ## some adjustments
# ## Use summary statistics (mean) to find and remove obvious  outlier in  sample 3
# #' means of all replicates of sample 3 
# dfstat <- df %>% filter(Sample,No==1) %>% group_by(SampleNo,RepNo) %>% summarise(mean=mean(RFU))
# dfstat
# #' Data that should be removed
# filter(df,SampleNo == 3 & RepNo == 3)
# #' I could not invert the filter though
# #' But it is the same as leaving out all the measurement from the well E01
# dfs <- filter(df, Well != "E06")
# #' all the measurements
# length(rownames(df))
# #'after 12 datapoints were removed
# length(rownames(dfs))
```

### All data with description

```{r }
## LOESS regression,
p <- ggplot(data = df) + 
  aes(x = MeasTimeSec, y=RFU) +
  geom_smooth(aes(color=SampleNo)) +
  geom_point(aes(colour=SampleNo))

#' Panel of plots
p+facet_wrap(~SampleNo) +
  coord_cartesian(xlim = c(0,1000)) +
  labs(title = "RNAse activity test", x = "sec", y = "RFU") +
  scale_colour_discrete(name="Conditions",
                        breaks=c(1,2,3,4,5,6,7,8),
                        labels=c("1: oldJoe1",
                                 "2: oldJoe10",
                                 "3: kit1",
                                 "4: kit10",
                                 "5: new1",
                                 "6: new10",
                                 "7: air",
                                 "8: air"))
#' All conditions together
  p + labs(title = "RNAse activity test", x = "sec",   y = "RFU", color = "Legend Title\n") +
  coord_cartesian(xlim = c(0,1000)) +
  scale_colour_discrete(name="Conditions",
                      breaks=c(1,2,3,4,5,6,7,8),
                      labels=c("1: oldJoe1",
                                 "2: oldJoe10",
                                 "3: kit1",
                                 "4: kit10",
                                 "5: new1",
                                 "6: new10",
                                 "7: air",
                                 "8: air"))


```

The warning notifies us we have zoomed in the plot by putting limits on axes.

### Plots of subsets of selected conditions being tested

```{r subsets, include=TRUE, echo=TRUE, eval=FALSE}

## positive contorls vs some negative controls
p <- df[raw$SampleNo==1|df$SampleNo==2|df$SampleNo==3|df$SampleNo==4|df$SampleNo==5|df$SampleNo==6|df$SampleNo==7|df$SampleNo==7,]
## LOESS regression
p <- ggplot(data = p) +
  aes(x = MeasTimeSec, y=RFU) +
  geom_smooth(aes(color=SampleNo)) +
  geom_point(aes(colour=SampleNo))


p + labs(title = "RNAse activity test", x = "sec",   y = "RFU", color = "Legend Title\n") +
  coord_cartesian(xlim = c(0,1000)) +
  scale_colour_discrete(name="Conditions",
                      breaks=c(1,2,3,4,5,7,8),
                      labels=c("1: oldJoe1",
                                 "2: oldJoe10",
                                 "3: kit1",
                                 "4: kit10",
                                 "5: new1",
                                 "6: new10",
                                 "7: air",
                                 "8: air"))


  
    
```

## Conclutions
- All substrates work, even the old one.
- For a decent signal, higher concentration than recommended is needed.
- Reaction kinetics is interesting with benzoase, very steep in fist few minutes. Presumably low copy number oftarget RNA in a real sample will prevent this from happening. 
- Reactions in the replicates were started in the order 1,2,3,4. The small delay is likely the reason for the trend in some measurments. This is apparent in tests with 10 fold the recommended substrate concentration. 

### What's next:
Test buffer compatibility. 



-----------------  
-----------------



