---
title: "RNAseAlert assay 11-2: 'correct' Cas13 purification"
author: "Jan Sklenar"
date: "04/05/2020"
output: html_document
editor_options: 
chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Correct Cas13a purification.
Recently We found that an incorrect Cas protein was expressed, purified, and used in all the tests so far. At least it is an explanation of failure to detect any signal. 
This assay is with newly expressed and purified Cas13a.

## Protocol:
![Copy of the original Excel table.](data/protocol11-2.svg)

#### Original RNAaseAlert [kit manual](https://www.thermofisher.com/document-connect/document-connect.html?url=https%3A%2F%2Fassets.thermofisher.com%2FTFS-Assets%2FLSG%2Fmanuals%2Ffm_1964.pdf&title=Uk5hc2VBbGVydCZ0cmFkZTsgTGFiIFRlc3QgS2l0IEluc3RydWN0aW9uIE1hbnVhbA==)


```{r libraries, echo=FALSE, include=FALSE}
#library("plyr")
library("dplyr")
#library("tidyr")
library("ggplot2")
```

## Data preparation
### Read in the data and parse some sample names 
```{r data, echo=TRUE, include=TRUE}
## read in raw data
raw <- read.delim("data/200604_test_11-2_correctCas132020-06-04T190003.txt",stringsAsFactors = FALSE)
colnames(raw)

#' remove 'X' column
raw <- raw[,1:10]
#' change names
colnames(raw)[7]
colnames(raw)[7] <- "MeasTimeSec"

#' parse 'Sample' to obtain sample number 'SampleNo'
raw$Sample[1]
raw$Sample[57]
#substr(raw$Sample,1,regexpr(' ', raw$Sample)-1) 
raw$SampleNo <- substr(raw$Sample,1,regexpr(' ', raw$Sample)-1)

#' in the same way extract replicate number - 'RepNo'
#substr(raw$Sample,regexpr(' ', raw$Sample)+1, regexpr('/', raw$Sample)-1)
raw$RepNo <- substr(raw$Sample,regexpr(' ', raw$Sample)+1, regexpr('/', raw$Sample)-1)

#' remove white space from 'Well' sting
#substr(raw$Well, 2, regexpr('$', raw$Well))[1:5]
raw$Well<- substr(raw$Well, 2, regexpr('$', raw$Well))

#' convert SampleNo to factor in order it to sort nicely in ggplots
raw$SampleNo <- as.factor(raw$SampleNo)
#raw$SampleNo <- factor(raw$SampleNo, c(seq(1:7),"air")) 
raw$SampleNo <- factor(raw$SampleNo, seq(1:8)) 
unique(raw$SampleNo)
#' View(raw) 
#'str(raw)
```

### Corrections
Not this time.

```{r corrections, include=FALSE, echo=FALSE, eval=FALSE}
## ## Condition 1 and 2 are the same thing. Create new SampleComb where 2 is replace with 1

raw$SampleComb <- raw$SampleNo 
raw[raw$SampleComb==2,"SampleComb"] <- 1

unique(raw$SampleNo)
unique(raw$SampleComb)
#' Which means the control with +RNAse has 8 replicates, unlike all the other conditions that has only 4. 
```



### Subset of data for plotting
```{r cleaned data}
## Selected (cleaned) data, only columns needed for plotting
#' Well, Sample, SampleNo, RepNo, MeasTimeSec, RFU
colnames(raw)
df <- raw[,c(3,6,7,8,11,12)] 
head(df) 
#View(df)
str(df)
table(is.na(df))
```
## Plots
### All data
```{r plots, eval=FALSE, include=FALSE, echo=FALSE}
#' all the data
#plot(x=raw$MeasTimeSec,y=raw$RFU)

#' colour by samples
ggplot(data = df) +
  aes(x = MeasTimeSec, y=RFU) +
  geom_point(aes(color=SampleNo))
```


### Adjustments 
```{r final result, include=TRUE, eval=TRUE, echo=TRUE}
## Adjustments - remove outliers if necessary

#' Measurement of this sample have many outliers
#dfp <- df %>% filter(SampleNo ==7) %>%
ggplot(subset(df, SampleNo %in% c(7))) +
  aes(x = MeasTimeSec, y=RFU) +
  geom_point(aes(color=SampleNo))

## Use summary statistics (mean) to find and remove obvious  outliers 
#' means of all replicates of sample 3
dfstat <- df %>% filter(SampleNo==1|SampleNo==2|SampleNo==3|SampleNo==4|SampleNo==5|SampleNo==6|SampleNo==7|SampleNo==8) %>% group_by(SampleNo,RepNo) %>% summarise(mean=mean(RFU))
dfstat
#View(dfstat)

#' Data that should be removed
df %>% filter(SampleNo==7 & RepNo==3)
#filter(df,SampleNo == 7 & RepNo == 3)

#View(df)
#' I do not know how to invert the filter 
#' Though it is the same as leaving out all the measurement from the wells in the first column
#'dfs <- filter(df, Well != "A01" & Well != "B01" & Well != "C01"&  Well != "D01"& Well != "E01"& Well != "F01" & Well != "G01"& Well != "H01")
dfs <- filter(df, Well != "G03")
#' all the measurements
length(rownames(df))
#'removinf onw column should remove 8*60 datapoints
length(rownames(dfs))
#' There are some zeros (negative spikes) in Sample "7 1/3" 
#' Let's removethose  zeros here
dfs[which(dfs$Sample=="7 1/3"&dfs$RFU==0),]
dfs[which(dfs$Sample=="7 1/3"&dfs$RFU==0),] <- NA

length(rownames(dfs))
table(is.na(dfs))
dfs <- dfs[complete.cases(dfs),]
table(is.na(dfs))
#dfs[which(is.na(dfs)),]


#' Correct - replace the original data with reduced dataframe  
df <- dfs
table(is.na(df))
```

### Samples and conditions for plot legends


```{r }
## Description
name <- "Samples"
breaks <- as.character(seq(1:8))
str(breaks)
breaks
labels <-  c("1: (+)RNA-O, Cas13",
            "2: (-)RNA-O, Cas13",
            "3: (+)RNA-O, noCas13",
            "4: (-)RNA-O, noCas13",
            "5: (+)RNA-S, Cas13",
            "6: (-)RNA-S, Cas13",
            "7: (+), substrate only",
            "8: (-), air")
```

```{r}
## LOESS regression,
p <- ggplot(data = df) + 
  aes(x = MeasTimeSec, y=RFU) +
  geom_smooth(aes(color=SampleNo)) +
  geom_point(aes(colour=SampleNo))

#' Panel of plots
p+facet_wrap(~SampleNo) +
  #coord_cartesian(xlim = c(0,1000)) +
  labs(title = "RNAse activity test", x = "sec", y = "RFU") +
  scale_colour_discrete(name=name, breaks=breaks, labels=labels)

#' All conditions together
  p + labs(title = "RNAse activity test", x = "sec",   y = "RFU", color = "Legend Title\n") +
  #coord_cartesian(xlim = c(0,1000)) +
  scale_colour_discrete(name=name, breaks=breaks, labels=labels)


```


### Plots of subsets of samples and conditions

```{r subsets, include=TRUE, echo=TRUE, eval=TRUE}
## Crude Orf1ab
p1 <- df[df$SampleNo==1|df$SampleNo==2|df$SampleNo==3|df$SampleNo==4,]
## Crude S
p2 <- df[df$SampleNo==5|df$SampleNo==6,]
## StrepTrap Orf1ab
p3 <- df[df$SampleNo==7|df$SampleNo==8,]
## StrepTrap S
#p4 <- df[df$SampleNo==9|df$SampleNo==10,]
## CIEX Orf1ab
#p5 <- df[df$SampleNo==11|df$SampleNo==12,]
## CIEX S
#p6 <- df[df$SampleNo==13|df$SampleNo==14,]
## substrate and air
#p7 <- df[df$SampleNo==15|df$SampleNo=="air",]

myplots <- list("p1"=p1,"p2"=p2,"p3"=p3)
class(myplots)
str(myplots)


## LOESS regression
for(p in myplots){
p <- as.data.frame(p)

print(ggplot(data = p) +
  aes(x = MeasTimeSec, y=RFU) +
  geom_smooth(aes(color=SampleNo)) +
  geom_point(aes(colour=SampleNo)) +
  labs(title = "RNAse activity test", x = "sec",   y = "RFU", color = "Legend Title\n") +
  #coord_cartesian(xlim = c(0,1000)) +
  scale_colour_discrete(name=name, breaks=breaks, labels=labels))

}
  
    
```


### Notes:

- For this assay I used high target concentration 1ul of 0.1ng/ul, which is approximately 50pM (1E13 copies, MW 53000) in 40ul assay on 96 well plate.

## Conclusions

### GOOD
- Cas13 collateral RNAse activity was detected with positive target RNA Orf1ab, but not with negative target RNA Orf1ab. 
- Omission of Cas13 yields no detctable activity.

### NOT SO GOOD
- S RNA does not work.  
- Sensitivity of the plate reader is not great.

-----------------  
-----------------



