---
title: "Test 12: Fluorescein dilution serie"
author: "Jan Sklenar"
date: "09/06/2020"
output: html_document
editor_options: 
chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Fluorimeter instrument test with Fluorecein

For some time I have a suspiccion that our spectofluorimeter, Varioskan (Thermo) is not sensitive enough. We have to use 10ul in the assay, while published protocols suggest to use only 1ul. We measure 40 ul in 96 well plates instead of 20ul in 384 though. These difference do not intuitivelly acount for lack of a signal in low concentration range. For this reason, Fluorescein disodium salt to benchmark sensitivity is used here.
The chemical is of purity > 80%, which should be sufficient for rough estimate of sensitivity.

## Protocol:
1mM fluorescein disolved in water, further diluted to 100 times (5ul in 0.5ml) several times to low aM.

Dilution|1|2|3|4|5|6|7|8|
|:------|-|-|-|-|-|-|-|-|
|concentration|1mM|10uM|100nM|1nM|100pM|1pM|10fM|100aM|





```{r libraries, echo=FALSE, include=FALSE}
#library("plyr")
library("dplyr")
#library("tidyr")
library("ggplot2")
```

## Data preparation
### Read in the data and parse some sample names 
```{r data, echo=TRUE, include=TRUE}
## read in raw data
raw <- read.delim("data/200609_test_12_Fluorescein_topReading2020-06-09T184501.txt",stringsAsFactors = FALSE)
colnames(raw)
head(raw)

#' remove 'X' column
raw <- raw[,1:10]
#' change names
colnames(raw)[7]
colnames(raw)[7] <- "MeasTimeSec"

#' parse 'Sample' to obtain sample number 'SampleNo'
raw$Sample[1]
unique(raw$Sample)
unique(raw$Well)

#substr(raw$Sample,1,regexpr(' ', raw$Sample)-1) 
raw$SampleNo <- substr(raw$Sample,1,regexpr(' ', raw$Sample)-1)
raw
#' in the same way extract replicate number - 'RepNo'
#substr(raw$Sample,regexpr(' ', raw$Sample)+1, regexpr('/', raw$Sample)-1)
raw$RepNo <- substr(raw$Sample,regexpr(' ', raw$Sample)+1, regexpr('/', raw$Sample)-1)

#' remove white space from 'Well' sting
#substr(raw$Well, 2, regexpr('$', raw$Well))[1:5]
raw$Well<- substr(raw$Well, 2, regexpr('$', raw$Well))

#' convert SampleNo to factor in order it to sort nicely in ggplots
raw$SampleNo <- as.factor(raw$SampleNo)
#raw$SampleNo <- factor(raw$SampleNo, c(seq(1:7),"air")) 
raw$SampleNo <- factor(raw$SampleNo, seq(1:16)) 
unique(raw$SampleNo)
#' View(raw) 
str(raw)
```

### Corrections - unnecessary


```{r corrections, include=FALSE, echo=FALSE, eval=FALSE}
## ## Condition 1 and 2 are the same thing. Create new SampleComb where 2 is replace with 1

raw$SampleComb <- raw$SampleNo 
raw[raw$SampleComb==2,"SampleComb"] <- 1

unique(raw$SampleNo)
unique(raw$SampleComb)
#' Which means the control with +RNAse has 8 replicates, unlike all the other conditions that has only 4. 
```



### Subset of data for plotting
```{r cleaned data}
## Selected (cleaned) data, only columns needed for plotting
#' Well, Sample, SampleNo, RepNo, MeasTimeSec, RFU
colnames(raw)
df <- raw[,c(3,6,7,8,11,12)] 
head(df) 
#View(df)
str(df)
#' any problematic dAta
table(is.na(df))
df[!complete.cases(df),]

```
## Plots
### All data
```{r plots, eval=TRUE, echo=TRUE, include=TRUE}
#' all the data
#plot(x=raw$MeasTimeSec,y=raw$RFU)

#' colour by samples
ggplot(data = df) +
  aes(x = MeasTimeSec, y=RFU) +
  geom_point(aes(color=SampleNo))
```


### Adjustments - unnecessary
```{r include=FALSE, eval=FALSE, echo=FALSE}
## Adjustments - remove outliers if necessary

#' Measurement of this sample have many outliers
#dfp <- df %>% filter(SampleNo ==7) %>%
ggplot(subset(df, SampleNo %in% c(7))) +
  aes(x = MeasTimeSec, y=RFU) +
  geom_point(aes(color=SampleNo))

## Use summary statistics (mean) to find and remove obvious  outliers 
#' means of all replicates of sample 3
dfstat <- df %>% filter(SampleNo==1|SampleNo==2|SampleNo==3|SampleNo==4|SampleNo==5|SampleNo==6|SampleNo==7|SampleNo==8) %>% group_by(SampleNo,RepNo) %>% summarise(mean=mean(RFU))
dfstat
#View(dfstat)

#' Data that should be removed
df %>% filter(SampleNo==7 & RepNo==3)
#filter(df,SampleNo == 7 & RepNo == 3)

#View(df)
#' I do not know how to invert the filter 
#' Though it is the same as leaving out all the measurement from the wells in the first column
#'dfs <- filter(df, Well != "A01" & Well != "B01" & Well != "C01"&  Well != "D01"& Well != "E01"& Well != "F01" & Well != "G01"& Well != "H01")
dfs <- filter(df, Well != "G03")
#' all the measurements
length(rownames(df))
#'removinf onw column should remove 8*60 datapoints
length(rownames(dfs))
#' There are some zeros (negative spikes) in Sample "7 1/3" 
#' Let's removethose  zeros here
dfs[which(dfs$Sample=="7 1/3"&dfs$RFU==0),]
dfs[which(dfs$Sample=="7 1/3"&dfs$RFU==0),] <- NA

length(rownames(dfs))
table(is.na(dfs))
dfs <- dfs[complete.cases(dfs),]
table(is.na(dfs))
#dfs[which(is.na(dfs)),]


#' Correct - replace the original data with reduced dataframe  
df <- dfs
table(is.na(df))
```

### Samples and conditions for plot legends


```{r }
## Description
name <- "Samples"
breaks <- as.character(seq(1:16))
str(breaks)
breaks
labels <-  c("1: 40ul 1mM",
            "2: 200ul 1mM",
            "3: 40ul 10uM",
            "4: 200ul 10uM",
            "5: 40ul 100nM",
            "6: 200ul 100nM",
            "7: 40ul 1nM",
            "8: 200ul 1nM",
            "9: 40ul 10pM",
            "10: 200ul 10pM",
            "11: 40ul 100fM",
            "12: 200ul 100fM",
            "13: 40ul 1fM",
            "14: 200ul 1fM",
            "15: 40ul 10aM",
            "16: 200ul 10aM")
```

```{r, final result, include=FALSE, eval=FALSE, echo=FALSE}
## LOESS regression,
p <- ggplot(subset(df,SampleNo %in% c("1","2"))) +
  aes(x = MeasTimeSec, y=RFU) +
  geom_smooth(aes(color=SampleNo)) +
  geom_point(aes(colour=SampleNo))

#' Panel of plots
p+facet_wrap(~SampleNo) +
  #coord_cartesian(xlim = c(0,1000)) +
  labs(title = "RNAse activity - substrate test with benzoase", x = "sec", y = "RFU") +
  scale_colour_discrete(name=name, breaks=breaks, labels=labels)

#' All conditions together
  p + labs(title = "RNAse activity - substrate test with benzoase", x = "sec",   y = "RFU", color = "Legend Title\n") +
  #coord_cartesian(xlim = c(0,1000)) +
  scale_colour_discrete(name=name, breaks=breaks, labels=labels)


```


### Final plots of subsets, samples, and conditions

```{r Final subsets, include=TRUE, echo=TRUE, eval=TRUE}
## Let us start with all data
## Bar plot
ggplot(data=df) +
  aes(x=SampleNo, y=RFU) +
  geom_bar(stat="identity",color="navyblue", fill="navyblue") +
  scale_y_continuous() +
  scale_x_discrete(labels=labels[1:16]) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
labs(title= "Fluorescein dilution serie",
      subtitle = "purity  >80%",
     x="Dilutions",
     y= "RFU",
     caption = "Average of 2 replicates") +
  theme(plot.subtitle=element_text(size=10, hjust=0, face="italic", color="black"))



## Subsets of interests are determined by volume in wells 
# volume = 40ul
p1 <- df[df$SampleNo==1|df$SampleNo==3|df$SampleNo==5|df$SampleNo==7|df$SampleNo==9|df$SampleNo==11|df$SampleNo==13|df$SampleNo==15,]
# volume = 200ul
p2 <- df[df$SampleNo==2|df$SampleNo==4|df$SampleNo==6|df$SampleNo==8|df$SampleNo==10|df$SampleNo==12|df$SampleNo==14|df$SampleNo==16,]
# list
myplots <- list("p1"=p1,"p2"=p2)
class(myplots)
#str(myplots)

# averave
#p1 %>% group_by(SampleNo) %>% summarise(mean=mean(RFU))

## Plots of different volumes individually
## normal scale
for(p in myplots){
p <- as.data.frame(p)

print(ggplot(data = p) +
  aes(x = SampleNo, y=RFU) +
  geom_point(aes(colour=SampleNo)) +
  labs(title = "Fluorescein - range of dilutions ", x = "Dilutions",   y = "RFU", color = "Legend Title\n") +
scale_colour_discrete(name=name, breaks=breaks, labels=labels))

}

## log scale
for(p in myplots){
p <- as.data.frame(p)

print(ggplot(data = p) +
  aes(x = SampleNo, y=RFU) +
  geom_bar(stat="identity",color="navyblue", fill="skyblue") +
  scale_y_log10() +
  labs(title = "Fluorescein - range of dilutions ",subtitle = "purity 80%", x = "Dilutions",   y = "log10(RFU)", color = "Legend Title\n") +
scale_colour_discrete(name=name, breaks=breaks, labels=labels))

}

```


### Notes:

- Fluorescein dilution serie  
**High concentration:**  
1mM = 1mmol /l = 1umol/ml = 1nmol/ul  
 In absolute values:  
40ul => 40nmol  
**Detection limit:**  
- We can detect 100nM Fluorescein in larger volume  (200ul) , which is  20pmol detected.

## Conclusions

### :-(
- Detection limit is still too high.  

### What next?
- A different plate reader
- Bottom reading
- Fine adjustments
- Larger volume

-----------------  
-----------------



