---
title: "RNAseAlert assay 09-2: Cas13 purification"
author: "Jan Sklenar"
date: "18/05/2020"
output: html_document
editor_options: 
chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Cas13 activity test follwed through purification steps.
This includes replicates and more substrate than assay 09.

## Protocol:
![Copy of the original Excel table.](data/protocol9-2.svg)

#### Original RNAaseAlert [kit manual](https://www.thermofisher.com/document-connect/document-connect.html?url=https%3A%2F%2Fassets.thermofisher.com%2FTFS-Assets%2FLSG%2Fmanuals%2Ffm_1964.pdf&title=Uk5hc2VBbGVydCZ0cmFkZTsgTGFiIFRlc3QgS2l0IEluc3RydWN0aW9uIE1hbnVhbA==)


```{r libraries, echo=FALSE, include=FALSE}
#library("plyr")
library("dplyr")
#library("tidyr")
library("ggplot2")
```

## Data preparation
### Read in the data and parse some sample names 
```{r data, echo=TRUE, include=TRUE}
## read in raw data
raw <- read.delim("data/200518_Cas13_02_all2020-05-18T171020.txt",stringsAsFactors = FALSE)
colnames(raw)

#' remove 'X' column
raw <- raw[,1:10]
#' change names
colnames(raw)[7]
colnames(raw)[7] <- "MeasTimeSec"

#' parse 'Sample' to obtain sample number 'SampleNo'
raw$Sample[1]
raw$Sample[57]
#substr(raw$Sample,1,regexpr(' ', raw$Sample)-1) 
raw$SampleNo <- substr(raw$Sample,1,regexpr(' ', raw$Sample)-1)

#' in the same way extract replicate number - 'RepNo'
#substr(raw$Sample,regexpr(' ', raw$Sample)+1, regexpr('/', raw$Sample)-1)
raw$RepNo <- substr(raw$Sample,regexpr(' ', raw$Sample)+1, regexpr('/', raw$Sample)-1)

#' remove white space from 'Well' sting
#substr(raw$Well, 2, regexpr('$', raw$Well))[1:5]
raw$Well<- substr(raw$Well, 2, regexpr('$', raw$Well))

#' convert SampleNo to factor in order it to sort nicely in ggplots
raw$SampleNo <- as.factor(raw$SampleNo)
raw$SampleNo <- factor(raw$SampleNo, c(seq(1:15),"air")) 
unique(raw$SampleNo)
#' View(raw) 
#'str(raw)
```

### Corrections
Not this time.

```{r corrections, include=FALSE, echo=FALSE, eval=FALSE}
## ## Condition 1 and 2 are the same thing. Create new SampleComb where 2 is replace with 1

raw$SampleComb <- raw$SampleNo 
raw[raw$SampleComb==2,"SampleComb"] <- 1

unique(raw$SampleNo)
unique(raw$SampleComb)
#' Which means the control with +RNAse has 8 replicates, unlike all the other conditions that has only 4. 
```



### Subset of data for plotting
```{r cleaned data}
## Selected (cleaned) data, only columns needed for plotting
#' Well, Sample, SampleNo, RepNo, MeasTimeSec, RFU
colnames(raw)
df <- raw[,c(3,6,7,8,11,12)] 
head(df) 
#View(df)
str(df)
table(is.na(df))
```
## Plots
### All data
```{r plots, eval=FALSE, include=FALSE, echo=FALSE}
#' all the data
#plot(x=raw$MeasTimeSec,y=raw$RFU)

#' colour by samples
ggplot(data = df) +
  aes(x = MeasTimeSec, y=RFU) +
  geom_point(aes(color=SampleNo))
```



```{r final result, include=TRUE, eval=TRUE, echo=FALSE}
## Adjustments (only when needed)

## Use summary statistics (mean) to find and remove obvious  outliers 
#' means of all replicates of sample 3
dfstat <- df %>% filter(SampleNo==1|SampleNo==2|SampleNo==3|SampleNo==4|SampleNo==5|SampleNo==6|SampleNo==7|SampleNo==8) %>% group_by(SampleNo,RepNo) %>% summarise(mean=mean(RFU))
dfstat
#View(dfstat)

#' Data that should be removed
df %>% filter(SampleNo==1 & RepNo==1)
filter(df,SampleNo == 3 & RepNo == 1)

#View(df)
#' I do not know how to invert the filter 
#' Though it is the same as leaving out all the measurement from the wells in the first column
dfs <- filter(df, Well != "A01" & Well != "B01" & Well != "C01"&  Well != "D01"& Well != "E01"& Well != "F01" & Well != "G01"& Well != "H01")
#' all the measurements
length(rownames(df))
#'removinf onw column should remove 8*60 datapoints
length(rownames(dfs))
#' Correct - replace the original data with reduced dataframe  
df <- dfs
table(is.na(df))
```

### Samples and conditions for plot legends


```{r }
## Description
name <- "Samples"
breaks <- as.character(c(seq(1:15),"air"))
str(breaks)
breaks
labels <-  c("1: (+)RNA-O, crude, inhib",
            "2: (-)RNA-O, crude, inhib",
            "3: (+)RNA-O, crude",
            "4: (-)RNA-O, crude",
            "5: (+)RNA-S, crude",
            "6: (-)RNA-S, crude",
            "7: (+)RNA-O, StrTrap",
            "8: (-)RNA-O, StrTrap",
            "9: (+)RNA-S, StrTrap",
            "10: (-)RNA-S, StrTrap",
            "11: (+)RNA-O, CIEX",
            "12: (-)RNA-O, CIEX",
            "13: (+)RNA-S, CIEX",
            "14: (-)RNA-S, CIEX",
            "15: (+) reporter",
            "16: (-) air")
```

```{r}
## LOESS regression,
p <- ggplot(data = df) + 
  aes(x = MeasTimeSec, y=RFU) +
  geom_smooth(aes(color=SampleNo)) +
  geom_point(aes(colour=SampleNo))

#' Panel of plots
p+facet_wrap(~SampleNo) +
  #coord_cartesian(xlim = c(0,1000)) +
  labs(title = "RNAse activity test", x = "sec", y = "RFU") +
  scale_colour_discrete(name=name, breaks=breaks, labels=labels)

#' All conditions together
  p + labs(title = "RNAse activity test", x = "sec",   y = "RFU", color = "Legend Title\n") +
  #coord_cartesian(xlim = c(0,1000)) +
  scale_colour_discrete(name=name, breaks=breaks, labels=labels)


```


### Plots of subsets of samples and conditions

```{r subsets, include=TRUE, echo=TRUE, eval=TRUE}
## Crude Orf1ab
p1 <- df[df$SampleNo==1|df$SampleNo==2|df$SampleNo==3|df$SampleNo==4,]
## Crude S
p2 <- df[df$SampleNo==5|df$SampleNo==6,]
## StrepTrap Orf1ab
p3 <- df[df$SampleNo==7|df$SampleNo==8,]
## StrepTrap S
p4 <- df[df$SampleNo==9|df$SampleNo==10,]
## CIEX Orf1ab
p5 <- df[df$SampleNo==11|df$SampleNo==12,]
## CIEX S
p6 <- df[df$SampleNo==13|df$SampleNo==14,]
## substrate and air
p7 <- df[df$SampleNo==15|df$SampleNo=="air",]

myplots <- list("p1"=p1,"p2"=p2,"p3"=p3,"p4"=p4,"p5"=p5,"p6"=p6,"p7"=p7)
class(myplots)
str(myplots)


## LOESS regression
for(p in myplots){
p <- as.data.frame(p)

print(ggplot(data = p) +
  aes(x = MeasTimeSec, y=RFU) +
  geom_smooth(aes(color=SampleNo)) +
  geom_point(aes(colour=SampleNo)) +
  labs(title = "RNAse activity test", x = "sec",   y = "RFU", color = "Legend Title\n") +
  #coord_cartesian(xlim = c(0,1000)) +
  scale_colour_discrete(name=name, breaks=breaks, labels=labels))

}
  
    
```

## Conclutions
### Notes:
Replicate 1 from samples 1-8 had to be removed, likely bubble in the multichannel pipette.


### What's next...?



-----------------  
-----------------



