---
title: "RNAseAlert kit test 03"
author: "Jan Sklenar"
date: "01/04/2020"
output: html_document
editor_options: 
chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# What is causing suppression of fluorescense when Cas13 is added?
## Design

- Cas13/impurities had an effect on fluorescence.
- We need to test Cas13 with and without its buffer to see what component has bearing on RFU. 
- We need to test DEPC in HEPES again, the first test did not work.
- Phosphate in higher concentration and generally ionic strength have profound effect on Benzoase. We can try 20mM Phosphate buffer.

#### Chemicals used in the reaction,
added in the following order in replicates of four, in squares on 96 well plate:

1. 20mM HEPES ph 7.0 buffer
2. Benzoase(TM)nuclease, 25U/ml stock solution
3. DEPC (inhibitor), 0.1% in the reaction 
4. Imidazole (inhibitor quencher), 10mM in the reaction
5. Benzoase(TM)nuclease (fresh enzyme after the background has been inhibited)
6. Cas13 2mg/ml was diluted: 5ul/153ul in Storage buffer (Josie), which gives 10x concentrated protein solution. For this test it is diluted to give 1x, 0.2x, and  0.01X diluted prtoein and bufer. 
7. substrate (fluorescein-RNA conjugate), two tubes designed for one test are here use for all the test - diluted to use 5ul in 48 replicates.  

#### Test conditions (volumes in microliters per well)
![Copy of original table in Excel.](data/protocol3.svg)

#### Here is Thermo [kit manual](https://www.thermofisher.com/document-connect/document-connect.html?url=https%3A%2F%2Fassets.thermofisher.com%2FTFS-Assets%2FLSG%2Fmanuals%2Ffm_1964.pdf&title=Uk5hc2VBbGVydCZ0cmFkZTsgTGFiIFRlc3QgS2l0IEluc3RydWN0aW9uIE1hbnVhbA==)


```{r libraries, echo=FALSE, include=FALSE}
#library("plyr")
library("dplyr")
#library("tidyr")
library("ggplot2")
```

## Data preparation
### Read in data and parse some sample names 
```{r data, echo=FALSE, include=FALSE}
## read in raw data
raw <- read.delim("data/200420_RNAaseAlert_test3.txt",stringsAsFactors = FALSE)
colnames(raw)
str(raw)
head(raw)
#' remove column
raw <- raw[,1:10]
#' change names
colnames(raw)[7]
colnames(raw)[7] <- "MeasTimeSec"

#' parse 'Sample' to obtain sample number 'SampleNo'
raw$Sample[1]
substr(raw$Sample,1,regexpr(' ', raw$Sample)-1) 
raw$SampleNo <- substr(raw$Sample,1,regexpr(' ', raw$Sample)-1)

#' in the same way extract replicate number - 'RepNo'
substr(raw$Sample,regexpr(' ', raw$Sample)+1, regexpr('/', raw$Sample)-1)
raw$RepNo <- substr(raw$Sample,regexpr(' ', raw$Sample)+1, regexpr('/', raw$Sample)-1)

#' remove white space from 'Well' sting
substr(raw$Well, 2, regexpr('$', raw$Well))[1:5]
raw$Well<- substr(raw$Well, 2, regexpr('$', raw$Well))
head(raw) 
str(raw)
```

### Corrections
Errors happen, see the table above.
```{r corrections}
## ## Condition 1 and 2 are the same thing. Create new SampleComb where 2 is replace with 1

raw$SampleComb <- raw$SampleNo 
raw[raw$SampleComb==2,"SampleComb"] <- 1

unique(raw$SampleNo)
unique(raw$SampleComb)
#' Which means the control with +RNAse has 8 replicates, unlike all the other conditions that has only 4. 
```



### Subset of data for plotting
```{r cleaned data}
## show cleaned data with selected columns
#' Well, Sample, SampleNo, RepNo, MeasTimeSec, RFU
df <- raw[,c(3,6,11,12,7,8,13)] 
head(df) 
str(df)
```
## Plots
### All data
```{r plots}
#' all the data
#plot(x=raw$MeasTimeSec,y=raw$RFU)

#' colour labeled samples
ggplot(data = df) +
  aes(x = MeasTimeSec, y=RFU) +
  geom_point(aes(color=SampleNo))
```
### Adjustments (if needed)
```{r final result}
# ## some adjustments
# ## Use summary statistics (mean) to find and remove obvious  outlier in  sample 3
# #' means of all replicates of sample 3 
# dfstat <- df %>% filter(SampleNo==1) %>% group_by(SampleNo,RepNo) %>% summarise(mean=mean(RFU))
# dfstat
# #' Data that should be removed
# filter(df,SampleNo == 3 & RepNo == 3)
# #' I could not invert the filter though
# #' But it is the same as leaving out all the measurement from the well E01
# dfs <- filter(df, Well != "E06")
# #' all the measurements
# length(rownames(df))
# #'after 12 datapoints were removed
# length(rownames(dfs))
```
### All data with description
```{r }
## LOESS regression
p <- ggplot(data = df) +
  aes(x = MeasTimeSec, y=RFU) +
  geom_smooth(aes(color=SampleNo)) +
  geom_point(aes(colour=SampleNo))

  p + labs(title = "RNAseAlert Test (with fresh substrate)", x = "min",   y = "RFU", color = "Legend Title\n") +
  coord_cartesian(xlim = c(0,1500)) +
  scale_colour_discrete(name="Conditions",
                      breaks=c(1,2,3,4,5,6,7,8,9,10,11,12),
                      labels=c("1: HEPES+RNAse",
                               "2: HEPES+RNAse",
                               "3: HEPES+RNAse+DEPC",
                               "4: HEPES+RNAse+DEPC+RNAse",
                               "5: Cas 0.01x",
                               "6: Cas 0.2x",
                               "7: Cas 1x",
                               "8: SB buffer 0.01x",
                               "9: SB buffer 0.2x",
                               "10: SB buffer 1x",
                               "11: PB+RNAse",
                               "12: PB+RNAse+DEPC"))
```
Not very readable. Let us split it.

### Ploting subsets by conditions being tested
This chunk is not going to be very ellegant; The annotation of all plots differ and we compare the same control with other measuremetns several time. Facet_wrap would be more elegant, but it does not give me flexibility to group plots that is neede here.   
```{r subsets}
## HEPES_DEPC
p <- raw[raw$SampleComb==1|raw$SampleComb==3|raw$SampleComb==4,]
## LOESS regression
p <- ggplot(data = p) +
  aes(x = MeasTimeSec, y=RFU) +
  geom_smooth(aes(color=SampleComb)) +
  geom_point(aes(colour=SampleComb))

  p + labs(title = "RNAseAlert Test: HEPES DEPC", x = "min",   y = "RFU", color = "Legend Title\n") +
  coord_cartesian(xlim = c(0,1500)) +
  scale_colour_discrete(name="Conditions",
                      breaks=c(1,3,4),
                      labels=c("1-2: HEPES+RNAse",
                               "3: +DEPC",
                               "4: +DEPC+RNAse"))

## HEPES Cas13
p <- raw[raw$SampleComb==1|raw$SampleComb==5|raw$SampleComb==6|raw$SampleComb==7,]
## LOESS regression
p <- ggplot(data = p) +
  aes(x = MeasTimeSec, y=RFU) +
  geom_smooth(aes(color=SampleComb)) +
  geom_point(aes(colour=SampleComb))

  p + labs(title = "RNAseAlert Test: HEPES Cas13", x = "min",   y = "RFU", color = "Legend Title\n") +
  coord_cartesian(xlim = c(0,1500)) +
  scale_colour_discrete(name="Conditions",
                      breaks=c(1,5,6,7),
                      labels=c("1-2: HEPES+RNAse",
                               "5: +Cas 0.01x",
                               "6: +Cas 0.2x",
                               "7: +Cas 1x"))

## HEPES Cas13 buffer
p <- raw[raw$SampleComb==1|raw$SampleComb==8|raw$SampleComb==9|raw$SampleComb==10,]
## LOESS regression
p <- ggplot(data = p) +
  aes(x = MeasTimeSec, y=RFU) +
  geom_smooth(aes(color=SampleComb)) +
  geom_point(aes(colour=SampleComb))

  p + labs(title = "RNAseAlert Test: HEPES Cas13 buffer", x = "min",   y = "RFU", color = "Legend Title\n") +
  coord_cartesian(xlim = c(0,1500)) +
  scale_colour_discrete(name="Conditions",
                      breaks=c(1,8,9,10),
                      labels=c("1-2: HEPES+RNAse",
                               "8: +Cas 0.01x",
                               "9: +Cas 0.2x",
                               "10: +Cas 1x"))

## Phosphate buffer
p <- raw[raw$SampleComb==1|raw$SampleComb==11|raw$SampleComb==12,]
## LOESS regression
p <- ggplot(data = p) +
  aes(x = MeasTimeSec, y=RFU) +
  geom_smooth(aes(color=SampleComb)) +
  geom_point(aes(colour=SampleComb))

  p + labs(title = "RNAseAlert Test: Phosphate buffer", x = "min",   y = "RFU", color = "Legend Title\n") +
  coord_cartesian(xlim = c(0,1500)) +
  scale_colour_discrete(name="Conditions",
                      breaks=c(1,11,12),
                      labels=c("1-2: HEPES+RNAse",
                               "11: PB+RNAse",
                               "12: PB+RNAse+DEPC"))
  
    
```

## Conclutions
    
### Good:
- DEPC inhibition works.
- Phosphate buffer in lower concentration works, the acitvity is lower.

 
### Not very good:
- The reaction were too diluted by mistake. 
- It seems Cas13 storage buffer inhibits the signal. 

### What next:
- We need to test real situation, Benzoase activity is irelevant after all.
- Now we know the assay works and some limits, we need start measuring Cas13 activity. 

-----------------  
-----------------



